### 2017-04-24 _paper_output/paper_sparse_hawkes.tex
more conveneient paper ignoring

% Default to the notebook output style

% Inherit from the specified cell style.

\documentclass[11pt]{article}

    \usepackage[T1]{fontenc}
    % Nicer default font (+ math font) than Computer Modern for most use cases
    \usepackage{mathpazo}

    % Basic figure setup, for now with no caption control since it's done
    % automatically by Pandoc (which extracts ![](path) syntax from Markdown).
    \usepackage{graphicx}
    % We will generate all images so they have a width \maxwidth. This means
    % that they will get their normal width if they fit onto the page, but
    % are scaled down if they would overflow the margins.
    \makeatletter
    \def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
    \else\Gin@nat@width\fi}
    \makeatother
    \let\Oldincludegraphics\includegraphics
    % Set max figure width to be 80% of text width, for now hardcoded.
    \renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=.8\maxwidth]{#1}}
    % Ensure that by default, figures have no caption (until we provide a
    % proper Figure object with a Caption API and a way to capture that
    % in the conversion process - todo).
    \usepackage{caption}
    \DeclareCaptionLabelFormat{nolabel}{}
    \captionsetup{labelformat=nolabel}

    \usepackage{adjustbox} % Used to constrain images to a maximum size 
    \usepackage{xcolor} % Allow colors to be defined
    \usepackage{enumerate} % Needed for markdown enumerations to work
    \usepackage{geometry} % Used to adjust the document margins
    \usepackage{amsmath} % Equations
    \usepackage{amssymb} % Equations
    \usepackage{textcomp} % defines textquotesingle
    % Hack from http://tex.stackexchange.com/a/47451/13684:
    \AtBeginDocument{%
        \def\PYZsq{\textquotesingle}% Upright quotes in Pygmentized code
    }
    \usepackage{upquote} % Upright quotes for verbatim code
    \usepackage{eurosym} % defines \euro
    \usepackage[mathletters]{ucs} % Extended unicode (utf-8) support
    \usepackage[utf8x]{inputenc} % Allow utf-8 characters in the tex document
    \usepackage{fancyvrb} % verbatim replacement that allows latex
    \usepackage{grffile} % extends the file name processing of package graphics 
                         % to support a larger range 
    % The hyperref package gives us a pdf with properly built
    % internal navigation ('pdf bookmarks' for the table of contents,
    % internal cross-reference links, web links for URLs, etc.)
    \usepackage{hyperref}
    \usepackage{longtable} % longtable support required by pandoc >1.10
    \usepackage{booktabs}  % table support for pandoc > 1.12.2
    \usepackage[inline]{enumitem} % IRkernel/repr support (it uses the enumerate* environment)
    \usepackage[normalem]{ulem} % ulem is needed to support strikethroughs (\sout)
                                % normalem makes italics be italics, not underlines

    % Colors for the hyperref package
    \definecolor{urlcolor}{rgb}{0,.145,.698}
    \definecolor{linkcolor}{rgb}{.71,0.21,0.01}
    \definecolor{citecolor}{rgb}{.12,.54,.11}

    % ANSI colors
    \definecolor{ansi-black}{HTML}{3E424D}
    \definecolor{ansi-black-intense}{HTML}{282C36}
    \definecolor{ansi-red}{HTML}{E75C58}
    \definecolor{ansi-red-intense}{HTML}{B22B31}
    \definecolor{ansi-green}{HTML}{00A250}
    \definecolor{ansi-green-intense}{HTML}{007427}
    \definecolor{ansi-yellow}{HTML}{DDB62B}
    \definecolor{ansi-yellow-intense}{HTML}{B27D12}
    \definecolor{ansi-blue}{HTML}{208FFB}
    \definecolor{ansi-blue-intense}{HTML}{0065CA}
    \definecolor{ansi-magenta}{HTML}{D160C4}
    \definecolor{ansi-magenta-intense}{HTML}{A03196}
    \definecolor{ansi-cyan}{HTML}{60C6C8}
    \definecolor{ansi-cyan-intense}{HTML}{258F8F}
    \definecolor{ansi-white}{HTML}{C5C1B4}
    \definecolor{ansi-white-intense}{HTML}{A1A6B2}

    % commands and environments needed by pandoc snippets
    % extracted from the output of `pandoc -s`
    \providecommand{\tightlist}{%
      \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
    \DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
    % Add ',fontsize=\small' for more characters per line
    \newenvironment{Shaded}{}{}
    \newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{{#1}}}}
    \newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{{#1}}}
    \newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
    \newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
    \newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{{#1}}}
    \newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{{#1}}}}
    \newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{{#1}}}
    \newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
    \newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{{#1}}}
    \newcommand{\RegionMarkerTok}[1]{{#1}}
    \newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{{#1}}}}
    \newcommand{\NormalTok}[1]{{#1}}

    % Additional commands for more recent versions of Pandoc
    \newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.53,0.00,0.00}{{#1}}}
    \newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{{#1}}}
    \newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.73,0.40,0.53}{{#1}}}
    \newcommand{\ImportTok}[1]{{#1}}
    \newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.73,0.13,0.13}{\textit{{#1}}}}
    \newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}
    \newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}
    \newcommand{\VariableTok}[1]{\textcolor[rgb]{0.10,0.09,0.49}{{#1}}}
    \newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{{#1}}}}
    \newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.40,0.40,0.40}{{#1}}}
    \newcommand{\BuiltInTok}[1]{{#1}}
    \newcommand{\ExtensionTok}[1]{{#1}}
    \newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.74,0.48,0.00}{{#1}}}
    \newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.49,0.56,0.16}{{#1}}}
    \newcommand{\InformationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}
    \newcommand{\WarningTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{{#1}}}}}

    % Define a nice break command that doesn't care if a line doesn't already
    % exist.
    \def\br{\hspace*{\fill} \\* }
    % Math Jax compatability definitions
    \def\gt{>}
    \def\lt{<}
    % Document parameters

\title{Regularised parameter selection in the non-stationary Hawkes process}

\author{Daniel MacKinlay}

    % Pygments definitions

\makeatletter
\def\PY@reset{\let\PY@it=\relax \let\PY@bf=\relax%
    \let\PY@ul=\relax \let\PY@tc=\relax%
    \let\PY@bc=\relax \let\PY@ff=\relax}
\def\PY@tok#1{\csname PY@tok@#1\endcsname}
\def\PY@toks#1+{\ifx\relax#1\empty\else%
    \PY@tok{#1}\expandafter\PY@toks\fi}
\def\PY@do#1{\PY@bc{\PY@tc{\PY@ul{%
    \PY@it{\PY@bf{\PY@ff{#1}}}}}}}
\def\PY#1#2{\PY@reset\PY@toks#1+\relax+\PY@do{#2}}

\expandafter\def\csname PY@tok@s2\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@nn\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\expandafter\def\csname PY@tok@o\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@fm\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\expandafter\def\csname PY@tok@gt\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.27,0.87}{##1}}}
\expandafter\def\csname PY@tok@sb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@kt\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.69,0.00,0.25}{##1}}}
\expandafter\def\csname PY@tok@nd\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
\expandafter\def\csname PY@tok@sa\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@mo\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@m\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@nt\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@no\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.53,0.00,0.00}{##1}}}
\expandafter\def\csname PY@tok@mf\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@na\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.49,0.56,0.16}{##1}}}
\expandafter\def\csname PY@tok@ni\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.60,0.60,0.60}{##1}}}
\expandafter\def\csname PY@tok@nf\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\expandafter\def\csname PY@tok@w\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.73,0.73}{##1}}}
\expandafter\def\csname PY@tok@sd\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@vc\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@go\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.53,0.53,0.53}{##1}}}
\expandafter\def\csname PY@tok@gi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.63,0.00}{##1}}}
\expandafter\def\csname PY@tok@ch\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@nc\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,1.00}{##1}}}
\expandafter\def\csname PY@tok@ne\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.82,0.25,0.23}{##1}}}
\expandafter\def\csname PY@tok@s\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@gr\endcsname{\def\PY@tc##1{\textcolor[rgb]{1.00,0.00,0.00}{##1}}}
\expandafter\def\csname PY@tok@gu\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.50,0.00,0.50}{##1}}}
\expandafter\def\csname PY@tok@s1\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@dl\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@k\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@cm\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@si\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.53}{##1}}}
\expandafter\def\csname PY@tok@nv\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@gh\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
\expandafter\def\csname PY@tok@mb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@cp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.74,0.48,0.00}{##1}}}
\expandafter\def\csname PY@tok@il\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@cpf\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@ow\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.67,0.13,1.00}{##1}}}
\expandafter\def\csname PY@tok@c1\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@ge\endcsname{\let\PY@it=\textit}
\expandafter\def\csname PY@tok@sr\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.53}{##1}}}
\expandafter\def\csname PY@tok@c\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@se\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.73,0.40,0.13}{##1}}}
\expandafter\def\csname PY@tok@vm\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@kc\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@sh\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@kd\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@cs\endcsname{\let\PY@it=\textit\def\PY@tc##1{\textcolor[rgb]{0.25,0.50,0.50}{##1}}}
\expandafter\def\csname PY@tok@sc\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.73,0.13,0.13}{##1}}}
\expandafter\def\csname PY@tok@gs\endcsname{\let\PY@bf=\textbf}
\expandafter\def\csname PY@tok@mi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@sx\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@nb\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@nl\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.63,0.63,0.00}{##1}}}
\expandafter\def\csname PY@tok@err\endcsname{\def\PY@bc##1{\setlength{\fboxsep}{0pt}\fcolorbox[rgb]{1.00,0.00,0.00}{1,1,1}{\strut ##1}}}
\expandafter\def\csname PY@tok@vg\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@kr\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@bp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@mh\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.40,0.40,0.40}{##1}}}
\expandafter\def\csname PY@tok@ss\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@gd\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.63,0.00,0.00}{##1}}}
\expandafter\def\csname PY@tok@kp\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}
\expandafter\def\csname PY@tok@vi\endcsname{\def\PY@tc##1{\textcolor[rgb]{0.10,0.09,0.49}{##1}}}
\expandafter\def\csname PY@tok@gp\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.00,0.50}{##1}}}
\expandafter\def\csname PY@tok@kn\endcsname{\let\PY@bf=\textbf\def\PY@tc##1{\textcolor[rgb]{0.00,0.50,0.00}{##1}}}

\def\PYZbs{\char`\\}
\def\PYZus{\char`\_}
\def\PYZob{\char`\{}
\def\PYZcb{\char`\}}
\def\PYZca{\char`\^}
\def\PYZam{\char`\&}
\def\PYZlt{\char`\<}
\def\PYZgt{\char`\>}
\def\PYZsh{\char`\#}
\def\PYZpc{\char`\%}
\def\PYZdl{\char`\$}
\def\PYZhy{\char`\-}
\def\PYZsq{\char`\'}
\def\PYZdq{\char`\"}
\def\PYZti{\char`\~}
% for compatibility with earlier versions
\def\PYZat{@}
\def\PYZlb{[}
\def\PYZrb{]}
\makeatother

    % Exact colors from NB
    \definecolor{incolor}{rgb}{0.0, 0.0, 0.5}
    \definecolor{outcolor}{rgb}{0.545, 0.0, 0.0}

    % Prevent overflowing lines due to hard-to-break entities
    \sloppy 
    % Setup hyperref package
    \hypersetup{
      breaklinks=true,  % so long urls are correctly broken across lines
      colorlinks=true,
      urlcolor=urlcolor,
      linkcolor=linkcolor,
      citecolor=citecolor,
      }
    % Slightly bigger margins than the latex defaults

    \geometry{verbose,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}

    \begin{document}

    \maketitle

    \section{$\ell_1$-penalised non-stationary Hawkes branching
process}\label{ellux5f1-penalised-non-stationary-hawkes-branching-process}

    \subsection{Abstract}\label{abstract}

We are interested in the problem of simultaneous identification of the
parameters of a Hawkes point process and identifying a sparse driving
signal by means of $\ell_1$ penalised regression where, by assumption,
the driving noise has a different form. This is analaogous to a blind
deconvolution problem or a compressive sensing-style inverse problem,
but considerably more challenging; While the process model superfcially
resembles a linear system, there are few of the conveneient results that
woudl apply in a linear regression problem; and as a time series problem
it is difficult to choose our penalty parameters by standard
non-parametric means such as cross validation. The proposed method uses
a heuristically-motivated Akaike-style Information criterion
regularisation parameter selection, which we test through simulation.
The proposed method works well in practice.

    \subsection{Introduction}\label{introduction}

We define call a counting process $N$ indexed by $t$ a (TODO)
predictable Poisson point process if it has an associated
\emph{intensity} process $\lambda$, such that $\lambda$ is predictable
with respect to the filtration generated by $N$, i.e.
\[\lambda(\mathcal{F(t^-)}):=\lim_{h\to 0} \frac{\mathrm E\left(N(t,t+h)\right)}{h}.\]

The Hawkes process (Hawkes, 1971) specifically, is the linear,
self-exciting point process; that is, given a "background" rate
$\mu:\mathbb{R}^+\rightarrow\mathbb{R}^+,$ and a non-negative influence
kernel with positive support
$\phi:\mathbb{R}^+\rightarrow\mathbb{R}^+\cap\{0\}$, the Hawkes process
has the following expression for its intensity:

\[
\begin{aligned}
\lambda(\mathcal{F(t)}^-_{(-\infty,t)}) &= \mu(t) + \phi_\theta * N\\
&= \mu(t)  + \int_{-\infty}^{\infty}\phi_\theta(t-s)dN(s)\\
&= \mu(t) + \int_{-\infty}^{t}\phi_\theta(t-s)dN(s)\\
&= \mu(t) + \sum_{t_i\lt t}\phi_\theta(t-t_i)
\end{aligned}
\]

In this problem, we are given a signal realisation of the count process
$N$ on an interval $T=[t_0,t_N]$. We are interested in identifying, from
this data, two unknown functional parameters; both the background rate
function $\mu$ and the influence kernel $\phi$.

    One such assumption would be penalization :citeTikhonov1965,Hoerl1970
-{}-I apply an additional penalty to the log likelihood function,
penalizing particular values of the parameters.

For the penalized log likelihood for parameter $\theta$ and penalty
$\pi$, I write

\[L_\pi(\mathbf t, \theta'):= L_\pi(\mathbf t, \theta')- \pi P(\theta')\]

$P$ here is a non-negative functional which penalizes certain values of
the parameter vector, and $\pi\geq 0$ is the penalty weight
hyperparameter.

As before, the estimate is defined as the maximizer:

\[\hat{\theta}_\pi(\mathbf t) = \operatorname{argmax}_\theta L_\pi(\mathbf t;\theta)\]

For non-parametric extension to a parametric model, one usually
penalizes only the non-parametric extensions to the model, such that
they vanish as the penalty term increases. :citeGreen1987

\[\lim_{\pi\to\infty}\hat{\theta}_\pi'(\mathbf t) =\hat{\theta}(\mathbf t)\]

I follow this practice here, penalizing only the deviations of
$\omega(t)$ from the parametric estimate.

Hereafter, I will drop the prime from the augmented parameter vector and
simply write $\theta$.

Penalization is more frequently presented in the context of generalized
linear regression from i.id. samples, but it also fits within a
generalized Maximum Likelihood estimation theory.

Various gradient ascent algorithms are known to perform well for
:cite::Simon2011,Wu2008. Pragmatically, for performance reasons, I used
a mixed strategy. My algorithm attempts to update marginal estimates for
each $\omega_i$ parameter more rapidly first via Newton's method
:citeBattiti1992,Ozaki1979 and uses conjugate gradient descent for the
parametric estimates of the Hawkes parameters. If the updates steps are
small this seems to be stable. There are various tuning parameters to
such algorithms.

Many variants of pathwise regularization and exist, such as backwards
stepwise regularization, versions with random initialization, Least
Angle Regression, :citeEfron2004a selection with integrated cross
validation and so on. :citeFriedman2010

Many variants of pathwise regularization and exist, such as backwards
stepwise regularization, versions with random initialization, Least
Angle Regression, :citeEfron2004a selection with integrated cross
validation and so on. :citeFriedman2010 For this particular model I have
found few results giving me analytic guarantees, so I use heuristics and
simulation-based verification. I do not declare that this method "best"
in any sense, or that it will find the correct global maximum penalized
likelihood etc. The method, however, is simple enough to prototype
rapidly and, as it turns out, performs well on test data. Results will
be presented in the next chapter.

    We take $\mu$ to be an unknown sparse, non-random signal in a particular
sense. We will assume that, given a linear basis
$\{\psi_j \}_{1\leq j\leq M}$, that we may write the background rate as
\[\mu(t) = \mu_0 + \sum_{j=1}^M  \omega_j \psi_j(t).\] \emph{Sparsity}
here will imply that the coefficients $\omega_j$ will be mostly 0.

In practice, we choose
\[\psi_j(t) := \mathbb{I}_{[\tau_{j},\tau_{j+1})}(t)\] for some lattice
$\{\tau_j\},$ which will correspond to the assumption that "sparse"
triggers will have localised support in time; we are looking for short
"spikes".

Similarly, we will approximate $\phi$ in a different functional basis
${f}_{1\leq k \leq K}$, such that $\phi(s):= \sum_{k=1}^K f(s;k)$. This
choice is likely arbitrary, but many exist in the literature trading off
computational and statistical convenience, including Laguerre
polynomials (Ogata and Akaike, 1982), Exponential functions (Schoenberg
2015; Rambaldi et al, 2015), and step functions (Eichler, 2016). As a
convenient choice, we propose a basis of smooth unimodal non-negative
functions with a computationally convenient form and a single mode, the
convenient Maxwell distribution,
\[\phi_j(t):=\sqrt{\frac{2}{\pi}} \frac{t^2\exp\left(-\frac{t^2}{2\sigma_j^2}\right)}{\sigma_j^3}.\]

A motivation for these specific choices is a particular real-world
problem: modelling viral memes in social media. We suspect that online
popularity is governed by some mixture of organic social sharing,
modeled by the self-exciting $\phi$ component, mixed with intermittent
external boosts, such as when an internet meme is featured on an
external medium such as television.

    mean delay $\mu=2\tau \sqrt{\frac{2}{\pi}}$

mode $\sqrt{2}\tau$

scale $\sigma=\sqrt{\frac{\tau^2(3 \pi - 8)}{\pi}}$

We take fixed $M$-point lattice $\tilde{\tau}$ and $K$-fixed scales
$\tilde{\sigma}.$

The parameter vector to be estimated then is
$\theta:=(\mu, \tilde{\omega}, \tilde{\kappa}).$

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{paper_sparse_hawkes_files/paper_sparse_hawkes_10_0.pdf}
    \end{center}
    { \hspace*{\fill} \\}

    \subsubsection{Optimization}\label{optimization}

The non-penalised log-likelihood function for a given realisation of the
data is given

\[L_\pi(\mathbf t;\theta):=-\int_0^T\lambda_\theta(t)dt + \int_0^T\log \lambda_\theta(t) dN_t - \pi\|\boldsymbol \omega_\theta(t)\|_1\]

And using the method of maximum likelihood we estimate our parameter
vector by numerical optimisation of
\[\hat{\theta}_\pi(\mathbf t) = \operatorname{argmax}_\theta L_\pi(\mathbf t;\theta).\]

There is no closed-form solution for the optimum, but the derivatives of
this function are easy to calculate manually see e.g. (Ozaki 1979). Here
howevfer, we use automatic differentiation to make more complex
penalisation procedure conveneient to differentiate (Baydin and
Pearlmutter 2014).

We are able to use these to devise rapid gradient-pased optimisation
procedure.

    To remedy this, I use Sugiura's finite-sample-corrected version, the
\emph{AICc} (Sugiura, 1978).
\[{\mathrm {AIC}(L(X;\hat{\theta}))}=2d-2\ln(L)\]

\[{\mathrm {AICc}}:={\mathrm {AIC}}+{\frac {2d(d+1)}{N-d-1}}\]

    \begin{center}
    \adjustimage{max size={0.9\linewidth}{0.9\paperheight}}{paper_sparse_hawkes_files/paper_sparse_hawkes_13_0.pdf}
    \end{center}
    { \hspace*{\fill} \\}

    \section{Hyperparameter selection}\label{hyperparameter-selection}

    \subsubsection{Degrees of freedom}\label{degrees-of-freedom}

Whilst we might wish to select the regularisation penalties by cross
validation, cross validation is not straightforward for time series.

The traditional means of performing model selection in Hawkes-type
models e.g. Ogata (1983), uses \emph{degrees-of-freedom} penalties
(Efron 1986). Specifically, the \emph{AIC}, the Akaike Information
Criterion (Akaike 1981) is widely popular.

The AIC formula, for a model $M$ fit to a given data set $X$, for
estimated parameter vector $\hat{\theta^M}$ with log likelihood $L$ and
degrees of freedom $d^M$

\[\mathrm{AIC}(X,M)=2d^M-2L^M(X,\hat{\theta}^M)\]

The degrees of freedom for many models taken to be equal to the
cardinality of the parameter vector $\theta$.

Given two ML fitted models, $(M_1,\hat{\theta}^M_1)$ and
$(M_2\,\hat{\theta}^M_2)$, the difference in their AIC values is an
estimate of the relative Kullback-Leibler divergence of the inferred
measures $\hat{\mu_1},\hat{\mu_2}$ from the unknown true distribution,
$\mu$ i.e.

\[\mathrm{AIC}(X,M_1)-\mathrm{AIC}(X,M_2) \simeq D_\mathrm{KL}(\mu\|\hat{\mu_1}) - D_\mathrm{KL}(\mu\|\hat{\mu_2})\]

However, the AIC does not immediately solve the problem of
regularizition parameter selection in maximum likelihood, where the
regularisation parameter may be a real value smoothly interpolating many
possible models. One can calculate an AIC for penalised models, but this
require onerous Hessian matrix inversions and becomes costly in high
dimensions, and in any case does not generalise readily to non-smooth
penalties. (Konishi and kitagawa 2008)

We proceed heuristically, based on the arguments from Lasso-type
linear-model regularisation. There, the number of non-zero coefficients
fit under a particular value of the regularization penalty $\pi$ in an
$\ell_1$ regression model provides an unbiassed estimate of the
effective degrees of freedom of the model (Zou et al, 2007.)

    This suggests an adjusted likelihood parameter of the following form:

\[
\widehat{\mathrm{AICc}}(X, \hat{\theta}_\pi) = {\frac {2Nd(\hat{\theta}_\pi) }{N-d(\hat{\theta}_\pi)-1}} - 2\ln(L(X;\hat{\theta}_\pi), X))
\]

where $d_\pi$ counts the number of non-zero coefficients estimated for
given regularization coefficient $\pi$.

    Extensive grid search for optimsal parameters is expensive, so we select
both regularisation parameters using optimal sampling by Gaussian
process regression of the combined objective function.

(Snoek et al, 2012) using the package \texttt{scikit-optimize} (andreh7
et al, 2017) .

    \begin{Verbatim}[commandchars=\\\{\}]

        ---------------------------------------------------------------------------

        AttributeError                            Traceback (most recent call last)

        /usr/lib/python3.4/pickle.py in \_getattribute(obj, name, allow\_qualname)
        271         try:
    --> 272             obj = getattr(obj, subpath)
        273         except AttributeError:

        AttributeError: 'module' object has no attribute 'objective'

    During handling of the above exception, another exception occurred:

        AttributeError                            Traceback (most recent call last)

        <ipython-input-8-3150368f60c1> in <module>()
          1 from skopt import load
          2 
    ----> 3 res\_gp = load(join(FIT\_DIR, 'additive\_cvge'))

        /home/nfs/z3420853/Virtualenvs/basic34/lib/python3.4/site-packages/skopt/utils.py in load(filename, **kwargs)
        122         Reconstructed OptimizeResult instance.
        123     """
    --> 124     return load\_(filename, **kwargs)

        /home/nfs/z3420853/Virtualenvs/basic34/lib/python3.4/site-packages/sklearn/externals/joblib/numpy\_pickle.py in load(filename, mmap\_mode)
        573                     return load\_compatibility(fobj)
        574 
    --> 575                 obj = \_unpickle(fobj, filename, mmap\_mode)
        576 
        577     return obj

        /home/nfs/z3420853/Virtualenvs/basic34/lib/python3.4/site-packages/sklearn/externals/joblib/numpy\_pickle.py in \_unpickle(fobj, filename, mmap\_mode)
        505     obj = None
        506     try:
    --> 507         obj = unpickler.load()
        508         if unpickler.compat\_mode:
        509             warnings.warn("The file '\%s' has been generated with a "

        /usr/lib/python3.4/pickle.py in load(self)
       1036                     raise EOFError
       1037                 assert isinstance(key, bytes\_types)
    -> 1038                 dispatch[key[0]](self)
       1039         except \_Stop as stopinst:
       1040             return stopinst.value

        /usr/lib/python3.4/pickle.py in load\_global(self)
       1323         module = self.readline()[:-1].decode("utf-8")
       1324         name = self.readline()[:-1].decode("utf-8")
    -> 1325         klass = self.find\_class(module, name)
       1326         self.append(klass)
       1327     dispatch[GLOBAL[0]] = load\_global

        /usr/lib/python3.4/pickle.py in find\_class(self, module, name)
       1375         \_\_import\_\_(module, level=0)
       1376         return \_getattribute(sys.modules[module], name,
    -> 1377                              allow\_qualname=self.proto >= 4)
       1378 
       1379     def load\_reduce(self):

        /usr/lib/python3.4/pickle.py in \_getattribute(obj, name, allow\_qualname)
        273         except AttributeError:
        274             raise AttributeError("Can't get attribute \{!r\} on \{!r\}"
    --> 275                                  .format(name, obj))
        276     return obj
        277

        AttributeError: Can't get attribute 'objective' on <module '\_\_main\_\_'>

    \end{Verbatim}

    % Add a bibliography block to the postdoc

\bibliographystyle{unsrt}
\bibliography{refs}

    \end{document}